Predicting the PCG PSeudo-Random Number Generator In Practice

### 2020-05-15 article/main.tex
lattice en mieux

@@

\section{babairounding}

We call \(S't\) the truncated vector : \(S'[\ell,64+\ell]\).

We consider the case were we successfully guessed the 6 upper bits of each \(S_i\) and the \(\ell\) lower bits of \(S_0\). Then we know the \(6+\ell\) upper bits of \(S't\).

(à détailer ?)

We had define \(S'\) such that \(S'\) is in the lattice \(L(A,2^{128},n)\) and its \(\ell \) lower bits are 0. Then \(S't\) is in \(L(1,2^{64},n)\).

If we consider the vector \(T'\) computed in step (numerodelastepenquestion), we notice it is close to \(S't\) in the sense of they have exactly the same \(6+\ell\) upper bits. So we want to recover \(S't\) by solving an approach \(CVP(L(A,2^{64},n),T')\).

Now we need to prove that the \(pwet\) obtained with the Babai rounding is indeed \(S't\).

We denote \(r = rounding(YG^{-1}) \).

\begin{align*}
\lVert S't - pwet \rVert &= \lVert S't - rG \rVert \\
&= \lVert S't - (r-YG^{-1} + YG^{-1})G \rVert\\
&= \lVert S't - Y + (r-YG^{-1})G \rVert\\
&\leqslant \lVert S't - Y \rVert + \lVert(r-YG^{-1})\rVert \times \lVert G\rVert\\	
\end{align*}

By definition, \(r\) is the closest integer vector to \(YG^{-1}\). But as \(S't\) is an element of the lattice, \(S'tG^{-1}\) is an integer vector. Thus \(r-YG^{-1}\) is smaller than \(S'tG^{-1}-YG^{-1}\).

Hence :
\begin{align*}
\lVert S't - pwet \rVert &\leqslant \lVert S't - Y \rVert + \lVert(r-YG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert S't - Y \rVert + \lVert(S'tG^{-1}-YG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert S't - Y \rVert + \lVert(S't-Y)\rVert \times \lVert G^{-1} \rVert  \times \lVert G\rVert\\
& 	\leqslant \lVert S't - Y \rVert \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert )\\
\end{align*}

But, by construction of \(Y\), we have that \(\lVert S't - Y \rVert \leqslant \sqrt{n2^{2(58-l)}} \). So

\[ \lVert S't - pwet \rVert \leqslant \sqrt{n2^{2(58-l)}}\times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert )\]\\

Now all we need to do is find sets of parameters \((n,l)\) such that the following inequality is satisfied: 
\[\sqrt{n2^{2(58-\ell)}} \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert) \leqslant \lVert SVP(L(A,2^{64},n))\rVert \].

We notice that only the left term of the inequality depends on \(\ell\).

Hence, some sets of parameters that satisfies this inequality are :
\begin{itemize}
	\item \(n = 3\) and \(\ell \geqslant 19\)
	\item \(n = 4\) and \(\ell \geqslant 13\)
	\item \(n = 5\) and \(\ell \geqslant 11\)
	\item \(n = 6\) and \(\ell \geqslant 8\)
\end{itemize}

>>>>

\section{lattice geometric progression}

\subsection{The lattice}

One of the key point of the algorithm is, given a modulus \(M\) and a given common ration \(\alpha\), to find a modular geometric progression knowing only the most significant bits of its \(n\) first terms.

We call \(L(\alpha,M,n)\) the set of n-tuples \(U = (u_0,\dots u_{n-1})\) satisfying :

\[u_{i} = \alpha u_{i-1} \mod M \] for all \(i\) in \(\{1,\dots,n-1\}\).

We easily notice that, if \(U\) and \(V\) are in \(L(\alpha,M,n)\), so are \(-U\) and \(U+V\), hence \(L(\alpha,M,n)\) is a lattice.

The lattice \(L(\alpha,M,n)\) is spawned by the lines of the following \(n\times n\) matrix :

\[\begin{pmatrix}
1&\alpha&\alpha^2&\dots&\alpha^{n-1}\\
0&M&0&\dots&0\\
0&0&M&\dots&0\\
\dots&\dots&\dots&\dots&\dots\\
0&0&0&\dots&M\\
\end{pmatrix}\]

We want to find an element \(U\) in the lattice only knowing the most significant bits of each of its coordinates.

We decompose \(U = T + (T-U)\) where \(T\) contains only the upper bits and \(T-U\) the lowers one. There is no reason for \(T\) to be in the lattice, but as \(T-U\) is small, we know \(T\) is \emph{close} to \(U\). So, to find \(U\), we will solve CVP(\(L(\alpha,M,n)\),\(T\)).

\subsection{A first way to solve the CVP : the Babai's rounding}

We compute \(G\) the LLL-reduction of the generating matrix of \(L(\alpha,M,n)\).

As \(T\) is not an element of the lattice, \(TG^{-1}\) is not an integer vector.

We fix \(r = rounding(TG^{-1}) \), then \(rG\) is an element of the lattice. What can we say about the distance between \(rG\) and \(U\) ?

\begin{align*}
\lVert U - rG \rVert &= \lVert U - (r-TG^{-1} + TG^{-1})G \rVert\\
&= \lVert U - T + (r-TG^{-1})G \rVert\\
&\leqslant \lVert U - T \rVert + \lVert(r-TG^{-1})\rVert \times \lVert G\rVert\\	
\end{align*}

By definition, \(r\) is the closest integer vector to \(TG^{-1}\). But as \(U\) is an element of the lattice, \(UG^{-1}\) is an integer vector. Thus \(r-TG^{-1}\) is smaller than \(UG^{-1}-TG^{-1}\).

Hence :
\begin{align*}
\lVert U - rG \rVert &\leqslant \lVert U - T \rVert + \lVert(r-TG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert U - T \rVert + \lVert(UG^{-1}-TG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert U - T \rVert + \lVert(U-Y)\rVert \times \lVert G^{-1} \rVert  \times \lVert G\rVert\\
& 	\leqslant \lVert U - T \rVert \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert )\\
\end{align*}

If we can prove that the left side of the inequality is smaller than the SVP of G, then wa have proved that \(rG\) is indeed the vector we were searching for.

@@

\subsection{preuve cas facile}

The vector \(S'\) has been constructed such that it is in \(L(A,2^{128},n)\) and its \(\ell\) lower bits are zero.

We call \(U\) the truncated vector : \(S'[\ell,64+\ell]\). We notice that \(U\) is in \(L(A,2^{64},n)\).

We consider the case were we successfully guessed the 6 upper bits of each \(S_i\) and the \(\ell\) lower bits of \(S_0\). Then we know the \(6+\ell\) upper bits of \(U\).

(maybe dessin)

We consider the vector \(T'\) computed in step (numerodelastepenquestion). To recover \(U\), we use the Babai's rounding explain in section (insererlasection) on \(T'\) and \(L(A,2^{64},n)\).

As \(U\) and \(T'\) have exactly the same \(6+\ell\) upper bits  know that \(\lVert U-T \rVert \leqslant \sqrt{n2^{2(58-l)}}\).

So, to be sure that our algorithm return the vector \(U\), we need to find sets of parameters \((n,\ell)\) such that :

\[\sqrt{n2^{2(58-\ell)}} \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert) \leqslant \lVert SVP(L(A,2^{64},n))\rVert \]

\begin{itemize}
	\item For \(n=3\)

	\(G = \begin{pmatrix}
	-1241281756092&3827459685972&-728312298332\\
	-5001120657083&-2117155768935&5479732607037\\
	8655886039732&3303731088004&6319848582548\\
	\end{pmatrix}\)

	\(\lVert G \rVert \times \lVert G^{-1} \rVert = 2.869065991049632\)

	\(\lVert SVP(G) \rVert = 4.08909120094543e{^12}\)

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 19\) %17 avec borne CVP

	\item For \(n=4\)

	\(G = \begin{pmatrix}
	-186304953996472&-126056243766680&7937589136904&93078431381544\\
	-216211368070119&99587582169277&-214303762177807&-1707551230219\\
	110964501361298&-5646098666150&-268280113597118&149382085707466\\
	131252974561432&-233919070109448&-98716819647784&-134620659538888\\
	\end{pmatrix}\)

	\(\lVert G \rVert \times \lVert G^{-1} \rVert = 2.060717188416003\)

	\(\lVert SVP(G) \rVert = 2.43569932844650e^{14}\)

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 13\) %11 avec borne CVP

	\item For \(n=5\)

	\(\lVert G \rVert \times \lVert G^{-1} \rVert = 3.7692142154392747\)

	\(\lVert SVP(G) \rVert = 1.71669732658884e^{15}\)

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 11\)  %9 avec borne CVP

	\item For \(n=6\)

	\(\lVert G \rVert \times \lVert G^{-1} \rVert = 2.6925925483728537\)

	\(\lVert SVP(G) \rVert =1.02550589786192e^{16}\)

	Hence, the smallest l such that the inequality is satisfied is \(\ell = 8\) %7 avec borne CVP
\end{itemize}
